A Large-Scale Study of the Use of Eval in R

### 2020-08-26 inst/paper/main.tex
Case study on base and environments

\begin{lstlisting}
> r <- 5
> env <- new.env()
> env$r <- 9
> eval(r, env)
[1] 5
> evalq(r, envir = env)
[1] 9
\end{lstlisting}

@@

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \c{eval}
can access the local scope and the global scope. In R, \c{eval} has an
argument \c{envir} to indicate in which environment its first argument,
after being resolved, must be evaluated.

By default, it is equal to \c{parent.frame()}, which is the parent
environment of the \c{eval} call.

@@

>>>>

\begin{lstlisting}
> r <- 5
> env <- new.env()
> env$r <- 9
> eval(r, envir = env)
[1] 5
> evalq(r, envir = env)
[1] 9
\end{lstlisting}

@@

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \c{eval}
can access the local scope and the global scope. In R, \c{eval} has an
argument \c{envir} to indicate in which environment its first argument,
after being resolved, must be evaluated. By default, it is equal to \c{parent.frame()}, which is the parent environment of the \c{eval} call.

 This argument can also be \c{NULL}, a list or a dataframe. In that case, \c{eval} copies \c{envir} and encloses it the environment defined by \c{enclos}. Finally, the argument can also be a number $n$. It means that the environment in which the expressions is evaluated will be the results of \c{sys.call(n)} where $n$ refers to the $n$-th call frame after the global environment.

\textbf{Do we speak more about what an environment is, \c{.GlobalEnv}, \c{emptyenv()}, environment navigation here, or in the Background section?}

We classify the environment from two points of view:
\begin{itemize}
	\item in which environment \c{eval} writes new bindings 
	\item in which \emph{existing} environment \c{eval} reads. It is the first non new environment. \c{eval} will also be able to read in all parent environment of that environment. 
\end{itemize}

% Variability: how many different envir class, and different envir expr, for one call site?

@@

There are 30 functions in \emph{base} that use \eval. Only five of them use \c{parse}, for legitimate reasons, such \c{source} and \c{sys.source} which load a file and then evaluate it, or \c{invokeRestartInteractively} which prompts the user for arguments to restart execution after an error.

We endeavored  to write these functions without \eval. \c{match.arg} appears in 60.1\% of the eval calls. It is used to get the arguments, possibly abbreviated, of a function. With one argument, as \c{match.arg(arg)}, it uses \c{eval} to infer them from a parent function's formal arguments. We can rewrite it using its version with two arguments, \c{match.arg(arg, choices)} where \c{choices} are the possible formal arguments of the function, as shown in Listing~\ref{code:matcharg}.

\begin{lstlisting}[caption={Rewriting \c{match.arg} without \eval}, label=code:matcharg]
f <- function(type=c("examples", "tests", "vignettes")) {
actual_type = match.arg(type)
...
}

f <- function(type=c("examples", "tests", "vignettes")) {
actual_type = match.arg(type, choices=c("examples", "tests", "vignettes"))
...
}
\end{lstlisting}
